---
title: "Chlorophyll variable name mystery"
author: 'E. Klein'
date: "2019-04-18"
output: 
  html_document:
    toc:  TRUE
    toc_float: TRUE
    theme: united
    highlight: tango
    code_folding: hide
editor_options: 
  chunk_output_type: console
---


```{r setup, cache = F, echo = F, message = F, warning = F, tidy = F}
# make this an external chunk that can be included in any file
require(knitr)
opts_chunk$set(echo =T, message = F, error = F, warning = F, comment = NA,  
               fig.align = 'left',  fig.width = 7.5, fig.height = 6,
               tidy = F, cache.path = '.cache/', fig.path = 'fig/')
options(width = 100, stringAsFactors=F, scipen=999)

library(RColorBrewer)
palette(brewer.pal(8, "Set2"))

library(ncdf4)
library(readr)
library(lubridate)
library(ggplot2)
library(dplyr)




```

Last run `r lubridate::now()`


# INTRODUCTION

In the CTD profiles and WQM burst average data collection you might encounter different names for the derived chlorophyll parameter. In a quick revision of the full collections three different names exists: 

- `CHLF`: Chlorophyll estimate from fluorescence readings using *factory* provided coefficients. In the ncdf metadata: "mass_concentration_of_inferred_chlorophyll_from_relative_fluorescence_units_in_sea_water,Concentration of inferred chlorophyll from relative fluorescence per unit volume of the water body,conversion from fluorescence to chlorophyll-a using factory coefficients,Milligrams per cubic metre,mg m-3"
- `CHLU`: Chlorophyll estimate from fluorescence readings using *user* provided coefficients. "mass_concentration_of_inferred_chlorophyll_from_relative_fluorescence_units_in_sea_water,Concentration of inferred chlorophyll from relative fluorescence per unit volume of the water body,conversion from fluorescence to chlorophyll-a using user coefficients,Milligrams per cubic metre,mg m-3"
- `CPHL`: Chlorophyll estimate from fluorescence readings using an *unknown* coefficient. In the ncdf metadata: "mass_concentration_of_inferred_chlorophyll_from_relative_fluorescence_units_in_sea_water,Concentration of inferred chlorophyll from relative fluorescence per unit volume of the water body,conversion from fluorescence to chlorophyll-a using unknown coefficients,Milligrams per cubic metre,mg m-3"

The idea is determine when and where those variables are present and to test if there are differences between the estimates when they are present in the same instrument.

# CTD  profiles

## Data Collection

The data is pulled from the geoserver service using the following query:

`http://geoserver-123.aodn.org.au/geoserver/ows?typeName=anmn_ctd_profiles_data&SERVICE=WFS&outputFormat=csv&REQUEST=GetFeature&VERSION=1.0.0*`

The metadata was pulled using the following query: 

`http://geoserver-123.aodn.org.au/geoserver/ows?typeName=anmn_ctd_profiles_map&SERVICE=WFS&outputFormat=csv&REQUEST=GetFeature&VERSION=1.0.0*`


The resulting table is stored locally so no need to repeat the geoserver query. 

```{r dataread}
## read CTD profiles from local file
CTD <- read_csv("IMOS_-_Australian_National_Mooring_Network_(ANMN)_-_CTD_Profiles.csv", skip=29)

## read metadata from the geoserver
CTD.map = read_csv("http://geoserver-123.aodn.org.au/geoserver/ows?typeName=anmn_ctd_profiles_map&SERVICE=WFS&outputFormat=csv&REQUEST=GetFeature&VERSION=1.0.0*")


## set chlorophyll variables as numeric
CTD$CHLF = as.numeric(CTD$CHLF)
CTD$CHLU = as.numeric(CTD$CHLU)
CTD$CPHL = as.numeric(CTD$CPHL)



```


The collection contains `r length(CTD$file_id)` CTD profiles from`r length(unique(CTD$cruise_id))` cruises


```{r CHLvalues}
## uncomment to re-run
## 
# CHL = CTD %>% group_by(file_id) %>%
#   summarise(CHLF = sum(!is.na(CHLF)),
#             CHLU = sum(!is.na(CHLU)),
#             CPHL = sum(!is.na(CPHL)))
# 
# 
# ## add metadata
# CHL = left_join(CHL, CTD.map)
# 
# ## get the name of the instrument from the thredds server
# ## as it is not in the .map file I need to get it from each individual file
# ## this operation could take ~5mins as there are ~1600 files to open
# urlPrefix = "http://thredds.aodn.org.au/thredds/dodsC/"
# instrument=character()
# 
# for (i in 1:nrow(CHL)){
#   print(i)
#   nc=nc_open(paste0(urlPrefix, CHL$url[i]))
#   instrument = append(instrument, ncatt_get(nc,0)$instrument)
#   nc_close(nc)
# }
# 
# CHL$instrument = instrument
# 
# write_csv(CHL, "CHL.csv")
```


```{r}
## read the metadata
## already processed file
CHL = read_csv("CHL.csv")

```


### Who is measuring what

A table of the instruments per chlorophyll variable

No instrument is reporting `CHLU`

No instrument is reporting more than one chlorophyll variable


```{r}
instrumentVariable = data.frame(Variable = c("CHLF", "CHLU", "CPHL"),
                                Instruments = c(paste0(unique(CHL$instrument[CHL$CHLF!=0]), collapse = ", "),
                                                paste0(unique(CHL$instrument[CHL$CHLU!=0]), collapse = ", "),
                                                paste0(unique(CHL$instrument[CHL$CPHL!=0]), collapse = ", ")))

knitr::kable(instrumentVariable)
```



## CHL variables by Station

This table shows the number of profiles that include each of the variable per `site_code`.

```{r}
siteCHL = CTD %>% group_by(site_code) %>% summarise(CHLF = length(unique(file_id[!is.na(CHLF)])), 
                                                    CPHL = length(unique(file_id[!is.na(CPHL)])))

knitr::kable(siteCHL)
```


# WQM timeseries

## Data collection

The data (burst averaged collection) is pulled from the geoserver service using the following query:

data: 
`http://geoserver-123.aodn.org.au/geoserver/ows?typeName=anmn_burst_avg_timeseries_data&SERVICE=WFS&outputFormat=csv&REQUEST=GetFeature&VERSION=1.0.0&`

metadata:
`http://geoserver-123.aodn.org.au/geoserver/ows?typeName=anmn_burst_avg_timeseries_map&SERVICE=WFS&outputFormat=csv&REQUEST=GetFeature&VERSION=1.0.0`

As the file are quite big, I will pull the data by `site_code`, one by one. Also see if there are concomitant measurements of two chlorophyll variables. The result table is stored for future needs.

```{r}

# get metadata first
WQM.map <- read_csv("http://geoserver-123.aodn.org.au/geoserver/ows?typeName=anmn_burst_avg_timeseries_map&SERVICE=WFS&outputFormat=csv&REQUEST=GetFeature&VERSION=1.0.")
## get the name of sites
sitenames = unique(WQM.map$site_code)


## 
## uncomment to re-run
##


# WQM.chl = data.frame(site_code = character(),
#                      CHLF = numeric(),
#                      CHLU = numeric(),
#                      CPHL = numeric(),
#                      CHLFCHLU = logical(),
#                      CHLFCPHL = logical(),
#                      CHLUCPHL = logical())
# 
# for (i in 1: length(sitenames)){
# 
#   print(paste0(i, " - ", sitenames[i]))
# 
#   GEOquery = paste0("http://geoserver-123.aodn.org.au/geoserver/ows?typeName=anmn_burst_avg_timeseries_data&SERVICE=WFS&outputFormat=csv&REQUEST=GetFeature&VERSION=1.0.0&CQL_FILTER=(site_code%20LIKE%20'", sitenames[i], "')")
#   WQM.data = suppressMessages(read_csv(GEOquery))
# 
# 
#   WQM.chl = bind_rows(WQM.chl, data.frame(site_code = sitenames[i],
#                                           CHLF = length(unique(WQM.data$timeseries_id[!is.na(WQM.data$CHLF)])),
#                                           CHLU = length(unique(WQM.data$timeseries_id[!is.na(WQM.data$CHLU)])),
#                                           CPHL = length(unique(WQM.data$timeseries_id[!is.na(WQM.data$CPHL)])),
#                                           CHLFCHLU = ifelse(sum(!is.na(WQM.data$CHLF) & !is.na(WQM.data$CHLU))>0,
#                                                             TRUE, FALSE),
#                                           CHLFCPHL = ifelse(sum(!is.na(WQM.data$CHLF) & !is.na(WQM.data$CPHL))>0,
#                                                             TRUE, FALSE),
#                                           CHLUCPHL = ifelse(sum(!is.na(WQM.data$CHLU) & !is.na(WQM.data$CPHL))>0,
#                                                             TRUE, FALSE)
#                                           ))
# 
# }
# 
# write_csv(WQM.chl, "WQMchl.csv")

WQM.chl = read_csv("WQMchl.csv")

knitr::kable(WQM.chl)
```


At PH100 and SYD100 have concomitant measurements of `CHLU` and `CHLF`. Let's explore the relationship between both.

### PH100

```{r}
i=12 ## PH100
GEOquery = paste0("http://geoserver-123.aodn.org.au/geoserver/ows?typeName=anmn_burst_avg_timeseries_data&SERVICE=WFS&outputFormat=csv&REQUEST=GetFeature&VERSION=1.0.0&CQL_FILTER=(site_code%20LIKE%20'", sitenames[i], "')")
WQM.data = suppressMessages(read_csv(GEOquery))

plot(WQM.data$CHLU, WQM.data$CHLF, pch=19, cex=0.5, col="blue")
model = lm(CHLU~CHLF, data=WQM.data)

CHLUCHLFbias.PH100 = mean(WQM.data$CHLU - WQM.data$CHLF, na.rm=T)

summary(model)
```

For PH100 there is a calibration factor of `r model$coefficients[2]` units of CHLF per unit of CHLU. The **mean bias** between `CHLU`and `CHLF` is **`r CHLUCHLFbias.PH100` mg/m続**.

### SYD100


```{r}
i=15 ## SYD100
GEOquery = paste0("http://geoserver-123.aodn.org.au/geoserver/ows?typeName=anmn_burst_avg_timeseries_data&SERVICE=WFS&outputFormat=csv&REQUEST=GetFeature&VERSION=1.0.0&CQL_FILTER=(site_code%20LIKE%20'", sitenames[i], "')")
WQM.data = suppressMessages(read_csv(GEOquery))

plot(WQM.data$CHLU, WQM.data$CHLF, pch=19, cex=0.5, col="blue")
model = lm(CHLU~CHLF, data=WQM.data)

CHLUCHLFbias.SYD100 = mean(WQM.data$CHLU - WQM.data$CHLF, na.rm=T)

summary(model)
```

For PH100 there is a calibration factor of `r model$coefficients[2]` units of CHLF per unit of CHLU. The **mean bias** between `CHLU`and `CHLF` is **`r CHLUCHLFbias.SYD100` mg/m続**.


# CONCLUSIONS 

Analysing the CTD profiles and the WQM instruments for the chlorophyll variables, this are the results of the exploration:

1. in the CTD profiles and the WQM burst average collections are three (3) different names for the chlorophyll, as separate variables: CHLF, CHLU, CPHL.
2. For the CTD collection: 
  + CHLU is not used anywhere in the CTD profiles. However it is used in 12 of the 28 WQM sites.
  + The name of the variable is associated with the instrument producing the measurement.
  + There are no cases with two CHL variables for the same profile exist
  + There are few stations (most of the NRS) that have alternate CHLF and CPHL variables in different profiles during its existence.
3. For the WQM burst average collection
  + All three variables could be found in the collection in different sites and different instruments
  + Only at PH100 and SYD100 exist concomitant measurements of `CHLU` and `CHLF`. The mean bias between `CHLU` and `CHLF` is `r CHLUCHLFbias.PH100` mg/m続 for PH100 and `r CHLUCHLFbias.SYD100` mg/m続 for SYD100 in 
  
  
# RECOMMENDATIONS

1. Add a new variable in each collection with the standard and unique name of `CHL` (and its corresponding burst statistics if applicable), and specify in the metadata the origin of the variable

2. Contact the BGC steering group to seek recommendation on what variable to use in case of the WQM register with `CHLU` and `CHLF` variables at the same timestamp.


  






